---
title: '**Spatial Economics -- Assignment 2**'
author: 
  - "Max Heinze (h11742049@s.wu.ac.at)"
  - "Jaime Miravet (h12235992@s.wu.ac.at)"
  - "Jan Trimmel (h11809096@s.wu.ac.at)"
date: "April 12, 2024"
output: 
  pdf_document:
    toc: true
    includes:
      in_header: !expr file.path(rprojroot::find_rstudio_root_file(), "helper", "wrap_code.tex")
header-includes: 
  - \usepackage{tcolorbox}
  - \usepackage{graphicx}
  - \usepackage[default]{lato}
papersize: a4
geometry: margin = 2cm
urlcolor: DarkOrchid!65!black
---

```{r, setup, include = FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 80), tidy = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
library(showtext)
font_add_google("Lato", "Lato")
showtext_auto()
```

\vspace{2em}

\begin{tcolorbox}
\centering \itshape The executable code that was used in compiling the assignment is available on GitHub at \url{https://github.com/maxmheinze/spatial}.
\end{tcolorbox}

\newpage

# Task A

Text

\newpage

# Task B

## Nice Maps

We create four maps:

```{r, out.width = "60%", fig.align = "center", echo = FALSE}

# Header ------------------------------------------------------------------

pacman::p_load(
  tidyverse,
  tmap,
  haven,
  foreign,
  magrittr,
  sf,
  stargazer,
  conleyreg
)


# Prepare Dataframe -------------------------------------------------------

litr <- read_dta("./assignment2/data/literacy_Arg-Bra-Par.dta", encoding = "ISO-8859-1")

shp0 <- read_sf("./assignment2/data/task_b_shapefile_0.shp")
shp1 <- read_sf("./assignment2/data/task_b_shapefile_1.shp")
shp2 <- read_sf("./assignment2/data/task_b_shapefile_2.shp")

spli <- shp2 %>%
  left_join(litr, by = c("NAME_2" = "muni"))


# Literacy Map ------------------------------------------------------------

tm_shape(spli) + 
  tm_fill("literacy", palette = "Reds", style = "equal", n = 10) + 
  tm_borders(col = "#BBBBBB") + 
  tm_shape(shp1) + 
  tm_borders(col = "#444444") + 
  tm_text("NAME_1", size = 0.75, col = "black", bg.color = "#FFFFFF") +
  tm_shape(shp0) + 
  tm_borders(col = "#000000") +
  tm_compass(position = c("right", "top"), size = 2) +
  tm_scale_bar(position = c("right", "bottom"), text.size = 0.5) +
  tm_layout(title = "Literacy Rate by Municipality", 
            title.bg.color = "white", 
            legend.position = c("left", "bottom"),
            legend.bg.color = "white", 
            frame = FALSE,
            fontfamily = "Lato")


# Population Density Map --------------------------------------------------

tm_shape(spli) + 
  tm_fill("popd", palette = "Blues", style = "quantile", n = 10) + 
  tm_borders(col = "#BBBBBB") + 
  tm_shape(shp1) + 
  tm_borders(col = "#444444") + 
  tm_text("NAME_1", size = 0.75, col = "black", bg.color = "#FFFFFF") +
  tm_shape(shp0) + 
  tm_borders(col = "#000000") +
  tm_compass(position = c("right", "top"), size = 2) +
  tm_scale_bar(position = c("right", "bottom"), text.size = 0.5) +
  tm_layout(title = "Population Density by Municipality", 
            title.bg.color = "white", 
            legend.position = c("left", "bottom"),
            legend.bg.color = "white", 
            frame = FALSE,
            fontfamily = "Lato")


# Jesuit Distance Map -----------------------------------------------------

tm_shape(spli) + 
  tm_fill("distmiss", palette = "Purples", style = "equal", n = 10) + 
  tm_borders(col = "#BBBBBB") + 
  tm_shape(shp1) + 
  tm_borders(col = "#444444") + 
  tm_text("NAME_1", size = 0.75, col = "black", bg.color = "#FFFFFF") +
  tm_shape(shp0) + 
  tm_borders(col = "#000000") +
  tm_compass(position = c("right", "top"), size = 2) +
  tm_scale_bar(position = c("right", "bottom"), text.size = 0.5) +
  tm_layout(title = "Distance from Jesuit Missions by Municipality", 
            title.bg.color = "white", 
            legend.position = c("left", "bottom"),
            legend.bg.color = "white", 
            frame = FALSE,
            fontfamily = "Lato")

# Franciscan Distance Map -------------------------------------------------

tm_shape(spli) + 
  tm_fill("distfran", palette = "Greens", style = "equal", n = 10) + 
  tm_borders(col = "#BBBBBB") + 
  tm_shape(shp1) + 
  tm_borders(col = "#444444") + 
  tm_text("NAME_1", size = 0.75, col = "black", bg.color = "#FFFFFF") +
  tm_shape(shp0) + 
  tm_borders(col = "#000000") +
  tm_compass(position = c("right", "top"), size = 2) +
  tm_scale_bar(position = c("right", "bottom"), text.size = 0.5) +
  tm_layout(title = "Distance from Franciscan Missions by Municipality", 
            title.bg.color = "white", 
            legend.position = c("left", "bottom"),
            legend.bg.color = "white", 
            frame = FALSE,
            fontfamily = "Lato")
```

## Replication of Table 2

```{r}

# Replicate Results -------------------------------------------------------

col1 <- lm(illiteracy ~ distmiss + lati + longi + corr + ita + mis + mis1, data = litr)
col2 <- lm(illiteracy ~ distmiss + lati + longi + area + tempe + alti + preci + rugg + river + coast + corr + ita + mis + mis1, data = litr)

litr_bra <- subset(litr, country == "BRA")
col3 <- lm(illiteracy ~ distmiss + lati + longi + as.factor(mesorregi), data = litr_bra)
col4 <- lm(illiteracy ~ distmiss + lati + longi + area + tempe + alti + preci + rugg + river + coast + as.factor(mesorregi), data = litr_bra)

litr_arg <- subset(litr, country == "Argentina")
col5 <- lm(illiteracy ~ distmiss + lati + longi + corr, data = litr_arg)
col6 <- lm(illiteracy ~ distmiss + lati + longi + area + tempe + alti + preci + rugg + river + coast + corr, data = litr_arg)

litr_pry <- subset(litr, country == "Paraguay")
col7 <- lm(illiteracy ~ distmiss + ita, data = litr_pry)
col8 <- lm(illiteracy ~ distmiss + area + tempe + alti + preci + rugg + river + coast + ita, data = litr_pry)
```

```{r, echo = FALSE, results='asis'}

cat("\\scalebox{.9}{")

stargazer(col1, col2, col3, col4, type = "latex", header=FALSE, no.space = TRUE, column.sep.width = "3pt", float = FALSE)

cat("}")
```

```{r, echo = FALSE, results='asis'}

cat("\\scalebox{.9}{")

stargazer(col5, col6, col7, col8, column.labels = c("(5)", "(6)", "(7)", "(8)"), model.numbers = FALSE, type = "latex", header=FALSE, no.space = TRUE, column.sep.width = "3pt", float = FALSE)

cat("}")
```


```{r}
lit1 <- litr %>%
  drop_na(lati, longi) %>%
  mutate(lat = lati,
         lon = longi)

col1c <- conleyreg(illiteracy ~ distmiss + lati + longi + corr + ita + mis + mis1, data = lit1,
           dist_cutoff = 11.112, lat = "lat", lon = "lon")
col2c <- conleyreg(illiteracy ~ distmiss + lati + longi + area + tempe + alti + preci + rugg + river + coast + corr + ita + mis + mis1, data = lit1,
           dist_cutoff = 11.112, lat = "lat", lon = "lon")

lit1_bra <- subset(lit1, country == "BRA")
col3c <- conleyreg(illiteracy ~ distmiss + lati + longi + as.factor(mesorregi), data = lit1_bra,
           dist_cutoff = 11.112, lat = "lat", lon = "lon")
col4c <- conleyreg(illiteracy ~ distmiss + lati + longi + area + tempe + alti + preci + rugg + river + coast + as.factor(mesorregi), data = lit1_bra,
           dist_cutoff = 11.112, lat = "lat", lon = "lon")

lit1_arg <- subset(lit1, country == "Argentina")
col5c <- conleyreg(illiteracy ~ distmiss + lati + longi + corr, data = lit1_arg,
           dist_cutoff = 11.112, lat = "lat", lon = "lon")
col6c <- conleyreg(illiteracy ~ distmiss + lati + longi + area + tempe + alti + preci + rugg + river + coast + corr, data = lit1_arg,
           dist_cutoff = 11.112, lat = "lat", lon = "lon")

lit1_pry <- subset(lit1, country == "Paraguay")
col7c <- conleyreg(illiteracy ~ distmiss + ita, data =  lit1_pry,
                   dist_cutoff = 11.112, lat = "lat", lon = "lon")
col8c <- conleyreg(illiteracy ~ distmiss + area + tempe + alti + preci + rugg + river + coast + ita, data =  lit1_pry,
                   dist_cutoff = 11.112, lat = "lat", lon = "lon")
```

```{r, echo = FALSE, results='asis'}

cat("\\scalebox{.9}{")

stargazer(col1c, col2c, col3c, col4c, type = "latex", header=FALSE, no.space = TRUE, column.sep.width = "3pt", float = FALSE)

cat("}")
```

```{r, echo = FALSE, results='asis'}

cat("\\scalebox{.9}{")

stargazer(col5c, col6c, col7c, col8c, column.labels = c("(5)", "(6)", "(7)", "(8)"), model.numbers = FALSE, type = "latex", header=FALSE, no.space = TRUE, column.sep.width = "3pt", float = FALSE)

cat("}")
```

\newpage

# Task C
## The perils of ignoring peer effects

\newpage

# Task D

The image is a screenshot, and those are conventionally stored in PNG format. The photo _contained_ in the screenshot is a photograph, and it is difficult to guess which format it was originally saved in. Let's say it's JPEG. Then, someone inserted the image into the assignment PDF, meaning it is technically not stored as a PNG anymore. What all those ways of storing the image have in common is that they are **raster formats**, as they consist of individual pixels. And even if we print the document containing the image, it gets printed as dots, which are not exactly pixels, but certainly form a raster rather than a vector.
