---
title: '**Spatial Economics -- Assignment 1**'
author: 
  - "Max Heinze (h11742049@s.wu.ac.at)"
  - "Kevin Kain (h12232066@s.wu.ac.at)"
  - "Jaime Miravet (h12235992@s.wu.ac.at)"
date: "April 2, 2024"
output: 
  pdf_document:
    toc: true
    includes:
      in_header: !expr file.path(rprojroot::find_rstudio_root_file(), "helper", "wrap_code.tex")
header-includes: 
  - \usepackage{tcolorbox}
  - \usepackage[default]{lato}
papersize: a4
geometry: margin = 2cm
urlcolor: DarkOrchid!65!black
---

```{r, setup, include = FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 80), tidy = TRUE)
```

\vspace{2em}

\begin{tcolorbox}
\centering \itshape The executable code that was used in compiling the assignment is available on GitHub at \url{https://github.com/maxmheinze/spatial}.
\end{tcolorbox}

\newpage

# Exercise 1

## Preliminaries

First, we load the `MASS` package and check what variables there are in the `Boston` dataset.

```{r, echo = TRUE, results = "hide"}
# Header ------------------------------------------------------------------

rm(list = ls())
gc()

pacman::p_load(MASS)

# Check Column Names ------------------------------------------------------

colnames(Boston)
```

## Creating the Function

Next, we create the desired function.

```{r, echo = TRUE}
# Create the function -----------------------------------------------------

boston_quick_ols <- function(dependent, ...) {
  
  # Create a formula string from the inputs
  independents <- paste(c(...), collapse = " + ")
  formula_string <- paste(dependent, "~", independents)
  
  # Fit the model 
  fitted_model <- lm(as.formula(formula_string), data = Boston)
  
  # Get the summary
  fitted_model_summary <- summary(fitted_model)
  
  # Get point estimates and confidence intervals
  list_coef <- fitted_model_summary$coefficients
  list_conf <- confint(fitted_model, level = 0.95)
  list_ervr <- fitted_model_summary$sigma^2
  
  # Output a list
  return(list(
    coefficients = list_coef[,1],
    error_variance = list_ervr,
    test_statistic_t = list_coef[,3],
    test_statistic_p = list_coef[,4],
    confidence_intervals = list_conf
  ))
  
}
```

## A Simple Linear Model

Next, we apply the function, using a collection of four independent variables.

```{r}
boston_quick_ols("medv", "rm", "age", "dis", "nox")
```

